---
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FuzzySimRes


The goal of FuzzySimRes, a library written in R, is to simulate synthetic fuzzy data and
provide so-called epistemic bootstrap procedures.

The simulation procedures are useful to generate synthetic samples that can be applied in statistical inference (see, e.g., [@grzegorzewskietal2020;@grzegorzewskiamcs2020]) and consist of fuzzy numbers (e.g., triangular or trapezoidal ones).

The epistemic bootstrap is used for epistemic fuzzy data (see [@Couso2014]) and it allows (see [@grzegorzewski2021;@Grzegorzewski_Romaniuk_2022]):

+ application of special resampling procedures, 
+ application of real-valued statistical tests for fuzzy data,
+ estimation of statistical characteristics (like the mean) of a fuzzy sample.

Additionally, the special epistemic fuzzy real-life data set is provided (see [@FARAZ20102684]) which consists of triangular fuzzy numbers.

The fuzzy data used in this package should be defined as the variables introduced in [@Gagolewski] (as, e.g., triangular fuzzy numbers).

The following procedures are available in the library:

+ Epistemic bootstrap procedures:
  + *EpistemicBootstrap* - the standard epistemic bootstrap,
  + *AntitheticBootstrap* - the antithetic epistemic bootstrap.
  
+ Random generation of the initial samples:  
  + *SimulateFuzzyNumber* - generation of a single fuzzy number (triangular, trapezoidal, PLFN one),
  + *SimulateSample* - generation of the whole fuzzy sample using various random distributions.
  
+ Applications of the epistemic bootstrap: 
  + *EpistemicEstimator* - estimation of the given statistical characteristic together with its SE/MSE using the fuzzy sample and epistemic bootstrap,
  + *EpistemicMean* - estimation of the mean using the fuzzy sample and epistemic bootstrap,
  + *EpistemicCorrectedVariance* - estimation of the variance with the special corrected formula using the fuzzy sample and epistemic bootstrap,
  + *EpistemicTest* - estimation of the p-value of the selected real-valued statistical test using the fuzzy sample and epistemic bootstrap,
  + *AverageStatisticEpistemicTest* - estimation of the p-value of the selected real-valued statistical test using the fuzzy sample, epistemic bootstrap, and averaging statistics method,
  + *MultiStatisticEpistemicTest* - estimation of the p-value of the selected real-valued statistical test using the fuzzy sample, epistemic bootstrap, and multi-statistics method,
  + *ResamplingStatisticEpistemicTest* - estimation of the p-value of the selected real-valued statistical test using the fuzzy sample, epistemic bootstrap, and resampling statistics method.
  


+ Additional procedures:
  + *CombinePValues* - combining the p-values of a statistical test into a single value,
  + *KSTestCriticalValue* - calculation of the p-value for a given critical value in the case of the one- and two-sample Kolmogorov-Smirnow test.
  
+ Dataset:
 + *controlChartData* - the real-life fuzzy epistemic data concerning electronic circuit thickness.


## Installation

You can install the latest development version of FuzzyResampling with:

``` r
library(devtools)
install_github("mroman-ibs/FuzzySimRes")
```


You can install the latest stable version from CRAN with:

``` r
install.packages("FuzzySimRes")
```

## Examples


``` {r}
# set seed

set.seed(12345)

# load library

library(FuzzySimRes)


# generate the initial sample consisting of 5 trapezoidal fuzzy numbers

sample1 <- SimulateSample(5,originalRandomDist="rnorm",parametersOriginalRD=list(mean=0,sd=1),
increasesCoreRandomDist="rexp", parametersCoreIncreasesRD=list(rate=2),
supportLeftRandomDist="runif",parametersSupportLeftRD=list(min=0,max=0.6),
supportRightRandomDist="runif", parametersSupportRightRD=list(min=0,max=0.6),
type="trapezoidal")

# apply the epistemic bootstrap with 10 cuts

EpistemicBootstrap (sample1$value, cutsNumber=10)

# apply the antithetic bootstrap with 10 cuts

AntitheticBootstrap (sample1$value, cutsNumber=10)

# calculate the mean using the standard epistemic bootstrap approach

EpistemicMean(sample1$value,cutsNumber = 30)

# calculate the median using the antithetic epistemic bootstrap approach

EpistemicEstimator(sample1$value,estimator="median",cutsNumber = 30,bootstrapMethod="anti")


# generate two independent initial fuzzy samples

list1<-SimulateSample(20,originalRandomDist="rnorm",parametersOriginalRD=list(mean=0,sd=1),
increasesCoreRandomDist="rexp", parametersCoreIncreasesRD=list(rate=2),
supportLeftRandomDist="runif",parametersSupportLeftRD=list(min=0,max=0.6),
supportRightRandomDist="runif", parametersSupportRightRD=list(min=0,max=0.6),
type="trapezoidal")


list2<-SimulateSample(20,originalRandomDist="rnorm",parametersOriginalRD=list(mean=0,sd=1),
increasesCoreRandomDist="rexp", parametersCoreIncreasesRD=list(rate=2),
supportLeftRandomDist="runif",parametersSupportLeftRD=list(min=0,max=0.6),
supportRightRandomDist="runif", parametersSupportRightRD=list(min=0,max=0.6),
type="trapezoidal")

# apply the Kolmogorov-Smirnov two-sample test for two different samples
# with the average statistics approach

EpistemicTest(list1$value,list2$value,cutsNumber = 30)

# apply the Kolmogorov-Smirnov two-sample test for two different samples
# with the multi-statistic and antithetic approach

EpistemicTest(list1$value,list2$value,algorithm = "ms",bootstrapMethod = "anti")

# apply the Kolmogorov-Smirnov one-sample test
# with the averaging statistic approach for the first sample

AverageStatisticEpistemicTest(list1$value,sample2=NULL,cutsNumber = 30,y="pnorm")

```

## References
